Reinforcement Learning for Options on Target Volatility Funds

### 2021-05-26 tex/targetvol.tex
Minor changes to the paper
Mostly grammar and phrasing fixes.

\documentclass[11pt]{article}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage[outer=3.75cm, marginparwidth=4.5cm]{geometry}
%\usepackage{marginnote}
\usepackage{amssymb,amsmath,amsthm,latexsym,amsfonts,amscd,dsfont,enumerate,bbm,mathabx}
\usepackage{graphicx,tikz}
\usepackage{draftwatermark}
\usepackage{enumerate} 
% \usepackage{listings}
% \usepackage{booktabs}
\usepackage{mathtools}
% \usepackage[gen]{eurosym}
% \usepackage{algorithm}
% \usepackage{algpseudocode}
\usepackage{soul}
\usepackage{comment}
\usepackage{bm}
\usepackage[font=small,labelfont=bf]{caption}
\graphicspath{{Local_Volatility/}{Black_Scholes}}
% \setbeamertemplate{bibliography item}[text]
% \renewcommand*{\bibfont}{\scriptsize}

@@

In this contribution we face a new topic never dealt with in the TVS literature: we deal with the funding costs coming from hedging the risky assets underlying the TVS. We consider the point of view of a bank selling a call option to a portfolio manager as protection on the capital invested in a TVS. The portfolio manager has the freedom of changing the relative weights of the risky assets during the life of the TVS. Since the risky assets have different hedging costs, the bank shall adjust the price of the protection to include them in the worst-case scenario, i.e. the most expensive strategy from the financing costs point of view. Hence, the pricing problem becomes a dynamical control problem over the risky portfolio composition. In our contribution we describe the dynamical control problem, and we derive an analytical solution in the simple case of the risky assets driven by a Black-Scholes (BS) model \citep{BlackScholes}. Then, we tackle the problem in the general case of a local volatility (LV) model \citep{Dupire94pricingwith,LVfirst} by using Reinforcement Learning (RL) algorithms, in particular the Proximal Policy Optimization developed in \textcite{ppo}.
\\

The paper is organized ad follows. In \Cref{sec:TVS} we describe the dynamics of a TVS in presence of valuation adjustments. Then, in \Cref{sec:Derivative} we study the case of derivative contracts on TVSs, we describe the dynamical control problem giving the expression of the Hamilton-Jacobi-Bellman equation and we derive the analytical results for plain vanilla options under BS dynamics. In \Cref{sec:RL} we illustrate how we have applied RL to solve the dynamic control problem, giving a deep description of the algorithm we have built. We conclude the paper with \Cref{sec:Numerical_results} where we present the numerical results obtained in this work for the Black and Scholes model and the local volatility (LV) one.

\section{Target Volatility Strategy}\label{sec:TVS}
In a TVS the fund manager selects an allocation strategy aiming at stabilizing the portfolio volatility to a target level. Clients investing in the fund pay a running fee for the service and their capital is protected. The fund manager usually buys an option on the TVS to ensure capital protection. For instance, the capital can be protected by buying a put option. In this case, we can write the net asset value (NAV) $A_t$ of the strategy as given by\footnote{Here we neglect discounting factors.}
\begin{equation}
    A_t \coloneqq \max\{V_t,K\} = V_t + (K-V_t)^+,
\end{equation}
where $V_t$ is the price process of the strategy, and $K$ is the guaranteed capital. On the other hand, the fund manager ca n replicate the payoff by means of the put-call parity by investing the capital in a low-risk asset and buying a call on the strategy
\begin{equation}
    A_t = K + (V_t-K)^+.
\end{equation}
In this way the TVS is only defined in the two contracts client-fund and fund-bank. The fund manager is not implementing the strategy by trading in the market, and he is not subject to additional costs to access the market. On the other way, the bank is paying such costs since she is actively hedging the call option sold to the manager.

@@

Now, in order to prevent arbitrages, we assume the existence of a risk-neutral measure $\mathbb{Q}$ under which the deflated gain processes of all traded security are martingales. Under this assumption we are able to derive the drift conditions on the security price processes, and in turn on the strategy price process.
\begin{equation}
    \forall T >t \quad \bar{G}_t^i = \mathbb{E}_t \left[\bar{G}_T^i\right] \quad\Longrightarrow \quad dS_t^i = r_tS_t^idt-d\pi_t^i-d\psi_t^i + dM_t^i,
\label{eq:risk_neutral}\end{equation}
where $M_t^i$ are martingale under $\mathbb{Q}$. If we substitute this expression for the security dynamics into the definition of the strategy we can check that the price process of the strategy is accruing at a cash account rate rate $r_t$ compensated for the fund manager fees
\begin{equation}
    dV_t = V_t(r_t-\phi_t)dt + dM_t^V,
\end{equation}
with $M_t^V$ martingale under $\mathbb{Q}$. Notice that,as expected from non-arbitrage considerations, the coupons paid by each security appear only in the drift of the security price process, but they do not impact the drift of the strategy.

@@

\begin{proposition}
When the underlying securities follow a Black and Scholes (BS) model with deterministic rates, the optimal strategy for a monotone payoff consists in minimizing the local drift function, independently of the current state $I_t$
\begin{equation}
    \alpha^*(t) \coloneqq \argmin_\alpha \frac{\alpha \cdot \mu(t)}{\|\alpha \cdot \nu(t) \|}.
\label{eq:BS_optimal_strategy}\end{equation}
\end{proposition}
The absence of stochastic elements in \Cref{eq:BS_optimal_strategy} makes the optimal strategy known \textit{a priori} with no simulation needed; in fact one can solve the optimization problem once for all $\forall t \in \mathcal{T}$ just looking at the market data $\mu(t)$ and $\nu(t)$ for the securities. Once $\alpha^*$ is known, then one can price the payoff by the following BS formula
\begin{equation}
	V_0^{\text{BS}} = BS(F^{\text{TVS}}(0,T;\alpha^*,K), K, T, \bar{\sigma}, D(0,T;\zeta)),
\label{eq:bs_closed_solution}\end{equation}
where $F^{\text{TVS}}(t,T;\alpha,K)$ is the TVS forward curve defined by
\begin{equation}
	F^{\text{TVS}}(t,T;\alpha, K) = I_t \exp\left[\int_t^T\left(r(u)-\ell_\alpha \left(u,K\right)\right)du\right],
\end{equation}
while $BS(F,K,T,\sigma,D)$ is the standard BS formula for an European option with forward curve $F$, strike $K$, time to maturity $T$, volatility $\sigma$ and discount factor $D$.  
\subsubsection{Black's Closed form solution}
In absence of constraints on the allocation strategy, we are able to derive a closed form solution to the BS problem \eqref{eq:BS_optimal_strategy}
\begin{lemma}
	Let be $\mu, \alpha \in \mathbb{R}^n$, $\nu \in \mathbb{R}^{n\times n}$ be a full rank matrix and $\Sigma\coloneqq \nu \tr{\nu}$. Then the closed solution of the optimization problem \eqref{eq:BS_optimal_strategy} is 
	\begin{equation}
			\alpha^* = - \frac{\Sigma^{-1} \cdot \mu}{\|(\Sigma^{-1} \cdot \mu)\cdot \nu\|}
	\end{equation}
\end{lemma}
\begin{proof}
Since the argument of the minimum \eqref{eq:BS_optimal_strategy} is zero-homogeneous, then we can rewrite the problem as
\begin{equation}
		\begin{aligned}
			&\text { minimize } \alpha \cdot \mu\\
			&\text { subject to } \|\alpha \cdot \nu\|^2=1
	\end{aligned}
\end{equation}
By setting the Lagrangian function associated with the problem
\begin{equation}
	\mathcal{L}\left(\alpha, \lambda\right)=\alpha \cdot \mu-\lambda\left(\|\alpha \cdot \nu\|^2-1\right) \, ,
\end{equation}
we obtain the first order conditions
\begin{eqsys}
	\frac{\partial \mathcal{L}}{\partial \alpha}=\mu -2\lambda \Sigma \cdot \alpha=0 \\
	\frac{\partial \mathcal{L}}{\partial \lambda}= \|\alpha \cdot \nu\|^2-1=0
	\, ,
\end{eqsys}
Then, by applying simple algebra, we obtain the analytical form of the free optimal strategy
\begin{equation}
	\alpha^* = \pm \frac{\Sigma^{-1} \cdot \mu}{\|(\Sigma^{-1} \cdot \mu)\cdot \nu\|}
\end{equation}
We take the minus sign to get the minimum value of the TVS local drift. 
\end{proof}
\subsubsection{Active Asset (or Bang Bang) Solution}
A closed form solution to the minimization of the local drift correction \eqref{eq:BS_optimal_strategy} can also be derived in the common case that all costs of carry are nonnegative and the only constraint on portfolio weights is nonnegativity, which would mean a long only strategy by the fund manager.

@@

\subsection{Hamilton-Jacobi-Bellman Equation for Target Volatility Options}\label{subsec:HJB}
In this section we want to provide to the reader a formal description of the dynamic problem associated to options on target volatility strategies by writing the Hamilton-Jacobi-Bellman equation for the derivative value. We prove that from this description it is available the same closed formula \eqref{eq:BS_optimal_strategy} for the time-dependent BS model derived from the Gy\"ongy.

In full generality we suppose that the risky securities follows a generic multidimensional diffusive process $X_t$ given by the following SDE
\begin{equation}
		dX_t = M(X_t)dt + \Sigma(X_t) \cdot dW_t.
\end{equation}
The TVS dynamics is given as follows 
\begin{equation}
	\frac{dI_t}{I_t} = \left(r_t(X_t) -  \frac{\bar{\sigma} \alpha_t}{\|\alpha_t \cdot \nu_t(X_t)\|}  \cdot \mu_t(X_t)    \right)dt +\frac{\bar{\sigma}\alpha_t }{\|\alpha_t \cdot \nu_t(X_t) \|} \cdot \nu_t(X_t) \cdot dW_t
\end{equation}
Given $X := X_t$ and $I := I_t$, we can write the HJB equation for $V := V(t,I,X)$ as follows
\begin{equation}
	\begin{aligned}
		\frac{\partial V}{\partial t} + \max_\alpha  \left\{   \left(r_t(X) - \bar{\sigma}\frac{\alpha \cdot \mu_t(X)}{\|\alpha \cdot \nu_t(X)\|}\right) I \frac{\partial V}{\partial I} + (\nabla_X V) \cdot M(X) + \frac{1}{2} \bar{\sigma}^2I^2 \frac{\partial^2V}{\partial I^2}  \right.\\ 
		\left. +\frac{1}{2}\Tr\left(\tr{\Sigma(X)}(H_XV)\Sigma(X)\right)  + (\nabla_{X,I}V) \cdot \Sigma(X) \cdot \left(I\bar{\sigma}\frac{\alpha \cdot \nu_t(X)}{\|\alpha \cdot \nu_t(X)\|}\right)\right\} = 0,
	\end{aligned}
\end{equation}
where $\Tr(A)$ is the trace operator of $A$, $\nabla_X V$ the gradient of $V$ w.r.t. $X$, $H_XV$ the Hessian matrix of $V$ w.r.t. $X$ and $\nabla_{X,I}V $ is the vector defined by:
\begin{equation}
	\nabla_{X,I} V= \tr{\left(\frac{\partial^2 V}{\partial X^1 \partial I},\dots, \frac{\partial^2 V}{\partial X^n \partial I} \right)}.
\end{equation}
We take out from the maximum operator all the elements that do not depend on the risky allocation strategy $\alpha$
\begin{equation}
	\begin{aligned}
		\frac{\partial V}{\partial t} +   r_t(X) I \frac{\partial V}{\partial I} + (\nabla_X V) \cdot M(X) + \frac{1}{2} \bar{\sigma}^2I^2 \frac{\partial^2V}{\partial I^2} + \frac{1}{2}\Tr\left(\tr{\Sigma(X)}(H_XV)\Sigma(X)\right)  \\ 
		+\max_\alpha  \left\{  -\frac{\partial V}{\partial I} I  \bar{\sigma}\frac{\alpha \cdot \mu_t(X)}{\|\alpha \cdot \nu_t(X)\|} + 
		(\nabla_{X,I}V) \cdot \Sigma(X) \cdot \left(I\bar{\sigma}\frac{\alpha \cdot \nu_t(X)}{\|\alpha \cdot \nu_t(X)\|}\right)\right\} = 0.
		\end{aligned}
\label{eq:HJB_general}\end{equation}
\Cref{eq:HJB_general} represent the Hamilton-Jacobi-Bellman equation describing the TVS dynamic problem for a generic dynamics of the risky securities underlying the portfolio.
In our work we have that\footnote{We use $\circ$ for the element-wise product between two vectors and between vector and matrix.} $X_t = S_t$, $M(X_t) \coloneqq (r_t\mathds{1} - \mu_t)\circ S_t$ and $\Sigma(X_t) = S_t \circ \nu_t$.

@@

In our work we adopt as learning strategy the proximal policy optimization (PPO) developed in \textcite{ppo} and \textcite{gae}. PPO is an actor-critic algorithm well-suited for continuous control problems. The term \virgolette{actor-critic} means that it estimates the value function  $J$, which represents the expected value of future rewards, and updates the policy in the direction suggested by $J$. In PPO the policy distribution and the value function are parameterized by a set of parameters $\theta$: $J^\theta$, $\pi^\theta$. The PPO collects a small batch of experiences interacting with the environment to update $\theta$ by computing an estimator of the policy gradient and plugging it into a stochastic gradient ascent (SGA) algorithm in order to maximize an objective function $L^{\text{PPO}}$. The PPO objective is defined as
\begin{equation}
	L^{\text{PPO}} = L^\pi - c_1 L^J + c_2 L^{E}
\end{equation}
where $L^\pi$ is the policy surrogate, $L^J$ the value function error term and $L^E$ the entropy bonus, while $c_1$ and $c_2$ are hyper-parameters of the algorithm. We refer to \textcite{ppo} for a deeper description of those terms. 
 We use the PPO implementation found in \textsc{OpenAI Baselines} \citep{baselines}, where the parameterization procedure is performed by using artificial neural networks (ANNs), where the set of parameters $\theta$ are the hidden weights of the networks. In the following section we describe the the way we have formalized the TVS problem in the Reinforcement Learning framework.

@@

\subsection{Policy and Action Parameterization}\label{subsec:policy_action}
As discussed before, PPO parameterizes the agent policy by means of an artificial neural network. In our case we use a feed forward neural network (FFNN), a network with an acyclic topology and completely characterized by the following hyper-parameters: number of hidden layers $N$, number of neurons per hidden layer $M$, activation function per hidden layer and output activation function. The number of output neurons is fixed by the action space dimension, which is the number of assets $n$ in the risky portfolio, while the number of  input neurons is determined by the dimensionality of the state \eqref{eq:state}, $n+2$. We work with a stochastic policy whose functional form is a multivariate diagonal Gaussian distribution where the mean $\mu^\theta(s_k)$ is the output vector of the FFNN and the log-standard deviation $\log\sigma$ is an external parameter
\begin{equation}
	\pi^\theta(s_k) \sim \mathcal{N}(\mu^\theta(s_k),e^{\log\sigma}).
\end{equation}
In the \textsc{OpenAI Baselines} implementation of PPO, the parameter $\log\sigma$ is state-independent, but it decreases as the number of PPO update iterations increases. The idea is that the log-standard deviation will be higher at the beginning of the training phase in order to guarantee a good exploration of the action space while it will be lower at the ending phase to avoid too much noise in the proximity of the optimal policy.

Another important topic is how to parameterize the agent action in order to satisfy some specific constraints on the risky asset allocation weights. In fact, if on one hand with free $\alpha$ it is sufficient to set bounds in the action space, on the other hand the presence of constraints on the risky strategy imposes the need to parameterize the output of the ANN.
In this work we study the constraint where at time $T_k\in\mathcal{T}$ the agent can invest in the risky market adopting long positions only:
\begin{equation}
	\alpha_{T_k} \cdot \mathds{1} = 1 \quad \alpha_{T_k}^i >0 \quad \forall i=1,\dots,n.
\end{equation} 
We study two different action parameterizations for this kind of constraint.
\begin{itemize}
	\item \textbf{$\bm{(n-1)}$-spherical angles}: we rewrite $\alpha_{T_k}^j = (x^i_k)^2$ and we use $(n-1)$-spherical coordinates with unitary radius. Thus the strategy is parameterized by a vector of angles $\phi_k = \tr{(\phi_k^1,\dots,\phi_k^{n-1})}$ as follows
	\begin{equation}
		\begin{aligned}
			\alpha_{T_k}^1 &\coloneqq \cos^2 (\phi_k^1), \\
			\alpha_{T_k}^2 &\coloneqq \sin^2(\phi_k^1)\cos^2(\phi_k^2),\\
			&\vdots \\
			\alpha_{T_k}^{n-1} &\coloneqq \sin^2(\phi_k^1)\dots\sin^2(\phi_k^{n-2})\cos^2(\phi_k^{n-1}),\\
			\alpha_{T_k}^n &\coloneqq \sin^2(\phi_k^1)\dots\sin^2(\phi_k^{n-2})\sin^2(\phi_k^{n-1}).
		\end{aligned}
		\label{eq:euler_angles}\end{equation}
	where $\phi_k^i\in[0,\pi/2]$ to avoid oscillations in $\alpha$ that could slow down the training. Thus, with this parameterization, the ANN sample a vector of angles ($a_k = \phi_k$) and thus the action space is reduced by one dimension.
	\item \textbf{Normalized action}: we restrict the agent actions to a subset of $(\mathbb{R}^+)^n$ and then normalize the action to obtain the corresponding risky allocation strategy through the formula
	\begin{equation}
		\alpha_{T_k} = \frac{a_k}{a_k \cdot \mathds{1}}.
		\label{eq:normalized_action}\end{equation}
\end{itemize}

>>>>

\documentclass[11pt]{article}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage[outer=3.75cm, marginparwidth=4.5cm]{geometry}
%\usepackage{marginnote}
\usepackage{amssymb,amsmath,amsthm,latexsym,amsfonts,amscd,dsfont,enumerate,bbm,mathabx}
\usepackage{graphicx,tikz}
\usepackage{draftwatermark}
\usepackage{enumerate} 
% \usepackage{listings}
% \usepackage{booktabs}
\usepackage{mathtools}
% \usepackage[gen]{eurosym}
% \usepackage{algorithm}
% \usepackage{algpseudocode}
\usepackage{soul}
\usepackage{comment}
\usepackage{bm}
\usepackage[font=small,labelfont=bf]{caption}
\graphicspath{{Local_Volatility/}{Black_Scholes/}}
% \setbeamertemplate{bibliography item}[text]
% \renewcommand*{\bibfont}{\scriptsize}

@@

In this contribution we face a new topic never dealt with in the TVS literature: the funding costs coming from hedging the risky assets underlying the TVS. We consider the point of view of a bank selling a call option to a portfolio manager as protection on the capital invested in a TVS. The portfolio manager has the freedom of changing the relative weights of the risky assets during the life of the TVS. Since the risky assets have different hedging costs, the bank shall adjust the price of the protection to include them in the worst-case scenario, i.e. the most expensive strategy from the financing costs point of view. Hence, the pricing problem becomes a dynamical control problem over the risky portfolio composition. In our contribution we describe the dynamical control problem, and we derive an analytical solution in the simple case of the risky assets driven by a Black-Scholes (BS) model \citep{BlackScholes}. Then, we tackle the problem in the general case of a local volatility (LV) model \citep{Dupire94pricingwith,LVfirst} by using Reinforcement Learning (RL) algorithms, in particular the Proximal Policy Optimization developed in \textcite{ppo}.
\\

The paper is organized ad follows. In \Cref{sec:TVS} we describe the dynamics of a TVS in presence of valuation adjustments. Then, in \Cref{sec:Derivative} we study the case of derivative contracts on TVSs, we describe the dynamical control problem giving the expression of the Hamilton-Jacobi-Bellman equation and we derive the analytical results for plain vanilla options under BS dynamics. In \Cref{sec:RL} we illustrate how we have applied RL to solve the dynamic control problem, giving a description of the algorithm we have built. We conclude the paper with \Cref{sec:Numerical_results} where we present the numerical results obtained in this work for the Black and Scholes model and the local volatility (LV) one.

\section{Target Volatility Strategy}\label{sec:TVS}
In a TVS the fund manager selects an allocation strategy aiming at stabilizing the portfolio volatility to a target level. Clients investing in the fund pay a running fee for the service and their capital is protected. The fund manager usually buys an option on the TVS to ensure capital protection. For instance, the capital can be protected by buying a put option. In this case, we can write the net asset value (NAV) $A_t$ of the strategy as given by\footnote{Here we neglect discounting factors.}
\begin{equation}
    A_t \coloneqq \max\{V_t,K\} = V_t + (K-V_t)^+,
\end{equation}
where $V_t$ is the price process of the strategy, and $K$ is the guaranteed capital. On the other hand, the fund manager can replicate the payoff by means of the put-call parity by investing the capital in a low-risk asset and buying a call on the strategy
\begin{equation}
    A_t = K + (V_t-K)^+.
\end{equation}
In this way the TVS is only defined in the two contracts client-fund and fund-bank. The fund manager is not implementing the strategy by trading in the market, and he is not subject to additional costs to access the market. On the other way, the bank is paying such costs since she is actively hedging the call option sold to the manager.

@@

Now, in order to prevent arbitrages, we assume the existence of a risk-neutral measure $\mathbb{Q}$ under which the deflated gain processes of all traded security are martingales. Under this assumption we are able to derive the drift conditions on the security price processes, and in turn on the strategy price process.
\begin{equation}
    \forall T >t \quad \bar{G}_t^i = \mathbb{E}_t \left[\bar{G}_T^i\right] \quad\Longrightarrow \quad dS_t^i = r_tS_t^idt-d\pi_t^i-d\psi_t^i + dM_t^i,
\label{eq:risk_neutral}\end{equation}
where $M_t^i$ are martingale under $\mathbb{Q}$. If we substitute this expression for the security dynamics into the definition of the strategy we can check that the price process of the strategy is accruing at a cash account rate rate $r_t$ compensated for the fund manager fees
\begin{equation}
    dV_t = V_t(r_t-\phi_t)dt + dM_t^V,
\end{equation}
with $M_t^V$ martingale under $\mathbb{Q}$. Notice that, as expected from non-arbitrage considerations, the coupons paid by each security appear only in the drift of the security price process, but they do not impact the drift of the strategy.

@@

\begin{proposition}
When the underlying securities follow a Black and Scholes (BS) model with deterministic rates, the optimal strategy for a non-decreasing European payoff consists in minimizing the local drift function, independently of the current state $I_t$
\begin{equation}
    \alpha^*(t) \coloneqq \argmin_\alpha \frac{\alpha \cdot \mu(t)}{\|\alpha \cdot \nu(t) \|}.
\label{eq:BS_optimal_strategy}
\end{equation}
Analogously, the optimal strategy for a non-increasing European payoff consists in maximizing the local drift function:
\begin{equation}
    \alpha^*(t) \coloneqq \argmax_\alpha \frac{\alpha \cdot \mu(t)}{\|\alpha \cdot \nu(t) \|}.
\label{eq:BS_optimal_strategy_put}
\end{equation}
\end{proposition}
The absence of stochastic elements in \Cref{eq:BS_optimal_strategy} makes the optimal strategy known \textit{a priori} with no simulation needed; in fact one can solve the optimization problem once for all $\forall t \in \mathcal{T}$ just looking at the market data $\mu(t)$ and $\nu(t)$ for the securities. Once $\alpha^*$ is known, then one can price the payoff by the following BS formula
\begin{equation}
	V_0^{\text{BS}} = BS(F^{\text{TVS}}(0,T;\alpha^*,K), K, T, \bar{\sigma}, D(0,T;\zeta)),
\label{eq:bs_closed_solution}\end{equation}
where $F^{\text{TVS}}(t,T;\alpha,K)$ is the TVS forward curve defined by
\begin{equation}
	F^{\text{TVS}}(t,T;\alpha, K) = I_t \exp\left[\int_t^T\left(r(u)-\ell_\alpha \left(u,K\right)\right)du\right],
\end{equation}
while $BS(F,K,T,\sigma,D)$ is the standard BS formula for an European option with forward curve $F$, strike $K$, time to maturity $T$, volatility $\sigma$ and discount factor $D$.  
\subsubsection{Black's Closed form solution}
In absence of constraints on the allocation strategy, we are able to derive a closed form solution to the BS problem \eqref{eq:BS_optimal_strategy}
\begin{lemma}
	Let be $\mu, \alpha \in \mathbb{R}^n$, $\nu \in \mathbb{R}^{n\times n}$ be a full rank matrix and $\Sigma\coloneqq \nu \tr{\nu}$. Then the closed solution of the optimization problem \eqref{eq:BS_optimal_strategy} is 
	\begin{equation}
			\alpha^* = - \frac{\Sigma^{-1} \cdot \mu}{\|(\Sigma^{-1} \cdot \mu)\cdot \nu\|}
	\end{equation}
\end{lemma}
\begin{proof}
Since the argument of the minimum \eqref{eq:BS_optimal_strategy} is zero-homogeneous, then we can rewrite the problem as
\begin{equation}
		\begin{aligned}
			&\text { minimize } \alpha \cdot \mu\\
			&\text { subject to } \|\alpha \cdot \nu\|^2=1
	\end{aligned}
\end{equation}
By setting the Lagrangian function associated with the problem
\begin{equation}
	\mathcal{L}\left(\alpha, \lambda\right)=\alpha \cdot \mu-\lambda\left(\|\alpha \cdot \nu\|^2-1\right) \, ,
\end{equation}
we obtain the first order conditions
\begin{eqsys}
	\frac{\partial \mathcal{L}}{\partial \alpha}=\mu -2\lambda \Sigma \cdot \alpha=0 \\
	\frac{\partial \mathcal{L}}{\partial \lambda}= \|\alpha \cdot \nu\|^2-1=0
	\, ,
\end{eqsys}
Then, by applying simple algebra, we obtain the analytical form of the free optimal strategy
\begin{equation}
	\alpha^* = \pm \frac{\Sigma^{-1} \cdot \mu}{\|(\Sigma^{-1} \cdot \mu)\cdot \nu\|}
\end{equation}
We take the minus sign to get the minimum value of the TVS local drift. 
\end{proof}
\subsubsection{Active Asset (or Bang Bang) Solution}
A closed form solution to the minimization of the local drift correction \eqref{eq:BS_optimal_strategy} can also be derived in the common case that all costs of carry are nonnegative and the only constraint on portfolio weights is nonnegativity, which would mean a long only strategy by the fund manager.

@@

\subsection{Hamilton-Jacobi-Bellman Equation for Target Volatility Options}\label{subsec:HJB}
In this section we want to provide to the reader a formal description of the dynamic problem associated to options on target volatility strategies by writing the Hamilton-Jacobi-Bellman equation for the derivative value. We prove that from this description one can recover the same closed formula \eqref{eq:BS_optimal_strategy} for the time-dependent BS model which was derived above from the Gy\"ongy Lemma.

In full generality we suppose that the risky securities follows a generic multidimensional diffusive process $X_t$ given by the following SDE
\begin{equation}
		dX_t = M(X_t)dt + \Sigma(X_t) \cdot dW_t.
\end{equation}
The TVS dynamics is given as follows 
\begin{equation}
	\frac{dI_t}{I_t} = \left(r_t(X_t) -  \frac{\bar{\sigma} \alpha_t}{\|\alpha_t \cdot \nu_t(X_t)\|}  \cdot \mu_t(X_t)    \right)dt +\frac{\bar{\sigma}\alpha_t }{\|\alpha_t \cdot \nu_t(X_t) \|} \cdot \nu_t(X_t) \cdot dW_t
\end{equation}
Given $X := X_t$ and $I := I_t$, we can write the HJB equation for $V := V(t,I,X)$ as follows
\begin{equation}
	\begin{aligned}
		\frac{\partial V}{\partial t} + \max_\alpha  \left\{   \left(r_t(X) - \bar{\sigma}\frac{\alpha \cdot \mu_t(X)}{\|\alpha \cdot \nu_t(X)\|}\right) I \frac{\partial V}{\partial I} + (\nabla_X V) \cdot M(X) + \frac{1}{2} \bar{\sigma}^2I^2 \frac{\partial^2V}{\partial I^2}  \right.\\ 
		\left. +\frac{1}{2}\Tr\left(\tr{\Sigma(X)}(H_XV)\Sigma(X)\right)  + (\nabla_{X,I}V) \cdot \Sigma(X) \cdot \left(I\bar{\sigma}\frac{\alpha \cdot \nu_t(X)}{\|\alpha \cdot \nu_t(X)\|}\right)\right\} = 0,
	\end{aligned}
\end{equation}
where $\Tr(A)$ is the trace operator of $A$, $\nabla_X V$ the gradient of $V$ w.r.t.~$X$, $H_XV$ the Hessian matrix of $V$ w.r.t.~$X$ and $\nabla_{X,I}V$ is the vector defined by:
\begin{equation}
	\nabla_{X,I} V= \tr{\left(\frac{\partial^2 V}{\partial X^1 \partial I},\dots, \frac{\partial^2 V}{\partial X^n \partial I} \right)}.
\end{equation}
We take out from the maximum operator all the elements that do not depend on the risky allocation strategy $\alpha$
\begin{equation}
	\begin{aligned}
		\frac{\partial V}{\partial t} +   r_t(X) I \frac{\partial V}{\partial I} + (\nabla_X V) \cdot M(X) + \frac{1}{2} \bar{\sigma}^2I^2 \frac{\partial^2V}{\partial I^2} + \frac{1}{2}\Tr\left(\tr{\Sigma(X)}(H_XV)\Sigma(X)\right)  \\ 
		+\max_\alpha  \left\{  -\frac{\partial V}{\partial I} I  \bar{\sigma}\frac{\alpha \cdot \mu_t(X)}{\|\alpha \cdot \nu_t(X)\|} + 
		(\nabla_{X,I}V) \cdot \Sigma(X) \cdot \left(I\bar{\sigma}\frac{\alpha \cdot \nu_t(X)}{\|\alpha \cdot \nu_t(X)\|}\right)\right\} = 0.
		\end{aligned}
\label{eq:HJB_general}\end{equation}
\Cref{eq:HJB_general} represent the Hamilton-Jacobi-Bellman equation describing the TVS dynamic problem for a generic dynamics of the risky securities underlying the portfolio.
In our work we have that\footnote{We use $\circ$ for the element-wise product between two vectors and between vector and matrix.} $X_t = S_t$, $M(X_t) \coloneqq (r_t\mathds{1} - \mu_t)\circ S_t$ and $\Sigma(X_t) = S_t \circ \nu_t$.

@@

In our work we adopt as learning strategy the proximal policy optimization (PPO) developed in \textcite{ppo} and \textcite{gae}. PPO is an actor-critic algorithm well-suited for continuous control problems. The term \virgolette{actor-critic} means that it estimates the value function  $J$, which represents the expected value of future rewards, and updates the policy in the direction suggested by $J$. In PPO the policy distribution and the value function are parameterized by a set of parameters $\theta$: $J^\theta$, $\pi^\theta$. The PPO collects a small batch of experiences interacting with the environment to update $\theta$ by computing an estimator of the policy gradient and plugging it into a stochastic gradient ascent (SGA) algorithm in order to maximize an objective function $L^{\text{PPO}}$. The PPO objective is defined as
\begin{equation}
	L^{\text{PPO}} = L^\pi - c_1 L^J + c_2 L^{E}
\end{equation}
where $L^\pi$ is a surrogate of the policy's average reward, $L^J$ the value function error term and $L^E$ the entropy bonus, while $c_1$ and $c_2$ are hyper-parameters of the algorithm. We refer to \textcite{ppo} for a deeper description of those terms. 
 We use the PPO implementation found in \textsc{OpenAI Baselines} \citep{baselines}, where the parameterization procedure is performed by using artificial neural networks (ANNs), and the set of parameters $\theta$ are the hidden weights of the networks. In the following section we describe  the way we have formalized the TVS problem in the Reinforcement Learning framework.

@@

\subsection{Policy and Action Parameterization}\label{subsec:policy_action}
As discussed before, PPO parameterizes the agent policy by means of an artificial neural network. In our case we use a feed forward neural network (FFNN), a network with an acyclic topology and completely characterized by the following hyper-parameters: number of hidden layers $N$, number of neurons per hidden layer $M$, activation function per hidden layer and output activation function. The number of output neurons is fixed by the action space dimension, which is the number of assets $n$ in the risky portfolio, while the number of  input neurons is determined by the dimensionality of the state \eqref{eq:state}, $n+2$. We work with a stochastic policy whose functional form is a multivariate diagonal Gaussian distribution where the mean $\mu^\theta(s_k)$ is the output vector of the FFNN and the log-standard deviation $\log\sigma$ is an external parameter
\begin{equation}
	\pi^\theta(s_k) \sim \mathcal{N}(\mu^\theta(s_k),e^{\log\sigma}).
\end{equation}
In the \textsc{OpenAI Baselines} implementation of PPO, the parameter $\log\sigma$ is state-independent, but it decreases as the number of PPO update iterations increases. The idea is that the log-standard deviation will be higher at the beginning of the training phase in order to guarantee a good exploration of the action space while it will be lower at the end to avoid too much noise in the proximity of the optimal policy.

Another important topic is how to parameterize the agent action in order to satisfy some specific constraints on the risky asset allocation weights. In fact, if on the one hand with free $\alpha$ it is sufficient to set bounds in the action space, on the other hand the presence of constraints on the risky strategy imposes the need to parameterize the output of the ANN.
In this work we study the constraint where at time $T_k\in\mathcal{T}$ the agent can invest in the risky market adopting long positions only:
\begin{equation}
	\alpha_{T_k} \cdot \mathds{1} = 1 \quad \alpha_{T_k}^i \geq 0 \quad \forall i=1,\dots,n.
\end{equation} 
We study two different action parameterizations for this kind of constraint.
\begin{itemize}
	\item \textbf{$\bm{(n-1)}$-spherical angles}: we rewrite $\alpha_{T_k}^j = (x^i_k)^2$ and we use $(n-1)$-spherical coordinates with unitary radius. Thus the strategy is parameterized by a vector of angles $\phi_k = \tr{(\phi_k^1,\dots,\phi_k^{n-1})}$ as follows
	\begin{equation}
		\begin{aligned}
			\alpha_{T_k}^1 &\coloneqq \cos^2 (\phi_k^1), \\
			\alpha_{T_k}^2 &\coloneqq \sin^2(\phi_k^1)\cos^2(\phi_k^2),\\
			&\vdots \\
			\alpha_{T_k}^{n-1} &\coloneqq \sin^2(\phi_k^1)\dots\sin^2(\phi_k^{n-2})\cos^2(\phi_k^{n-1}),\\
			\alpha_{T_k}^n &\coloneqq \sin^2(\phi_k^1)\dots\sin^2(\phi_k^{n-2})\sin^2(\phi_k^{n-1}).
		\end{aligned}
		\label{eq:euler_angles}\end{equation}
	where $\phi_k^i\in[0,\pi/2]$ to avoid oscillations in $\alpha$ that could slow down the training. Thus, with this parameterization, the ANN sample a vector of angles ($a_k = \phi_k$) and thus the action space is reduced by one dimension.
	\item \textbf{Normalized action}: we restrict the agent actions to a subset of $(\mathbb{R}^+)^n$ and then normalize the action to obtain the corresponding risky allocation strategy through the formula
	\begin{equation}
		\alpha_{T_k} = \frac{a_k}{a_k \cdot \mathds{1}}.
		\label{eq:normalized_action}\end{equation}
\end{itemize}
